---
title: "R Notebook"
output: html_notebook
---
Step 2:
Using data in step 1 to analyse difference between two donor type (progression and stable). Draw MA plot, plot counts, variance plot, heatmap, sample-to-sample heatmap, and PCA plot.

(1)load data in R for the DESeq2.
```{r}
rm(list = ls())
counts0<-read.csv("counts.csv",row.names = 1)
donor0<-read.csv("donor.csv",row.names = 1)
colnames(counts0) == donor0$icgc_donor_id
```
(2)load data in DESeq2 (donor type, donor sex, and donor age at diagnosis).
```{r}
library("DESeq2")
dds<-DESeqDataSetFromMatrix(countData = counts0,
                            colData = donor0,
                            design = ~donor_sex + donor_type + donor_age_at_diagnosis)
```
(3)Pre-filtering: keep the data have more then 10 values in a raw.
```{r}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```
(4)Note on factor levels.
```{r}
dds$donor_type <- factor(dds$donor_type, levels = c("progression","stable"))
```
(5) Differential expression analysis and save data
```{r}
dds<-DESeq(dds)
res <- results(dds, alpha=0.05)
write.csv(as.data.frame(res), 
          file="donor_type_progression_stable.csv")
```
```{r}
head(res)
```
(6)Log fold change shrinkage for visualization and ranking by LFC estimates.
```{r}
resultsNames(dds)
resLFC <- lfcShrink(dds, coef="donor_type_stable_vs_progression", type="apeglm")
sum(resLFC$padj < 0.05, na.rm=TRUE)
sum(resLFC$padj < 0.01, na.rm=TRUE)
```
(7) draw MS plot with apeglm, normal, and ashr.
```{r}
plotMA(res, ylim=c(-0.3,0.3))
resNorm <- lfcShrink(dds, coef=2, type="normal")
resAsh <- lfcShrink(dds, coef=2, type="ashr")
par(mfrow=c(1,3), mar=c(4,4,2,1))
xlim <- c(0.1,1e6); ylim <- c(-2,2)
plotMA(resLFC, xlim=xlim, ylim=ylim, main="apeglm")
plotMA(resNorm, xlim=xlim, ylim=ylim, main="normal")
plotMA(resAsh, xlim=xlim, ylim=ylim, main="ashr")
```
(8)draw MA-plot with resLFC because it removes the noise associated with log2 fold changes from low count genes without requiring arbitrary filtering thresholds.
```{r}
plotMA(resLFC, ylim=c(-2,2))
```
(9) Draw Plot counts with dds. Here I specify the gene which had the smallest p value from the results table created above. You can select the gene to plot by row name or by numeric index.
```{r}
d <- plotCounts(dds, gene=which.min(res$padj), intgroup="donor_type", 
                returnData=TRUE)
library("ggplot2")
ggplot(d, aes(x=donor_type, y=count)) + 
  geom_point(position=position_jitter(w=0.1,h=0)) + 
  scale_y_log10(breaks=c(50,500,5000))
```
(10)Extracting transformed values
```{r}
vsd <- vst(dds, blind=FALSE)
ntd <- normTransform(dds)
```
(11) Effects of transformations on the variance
```{r}
library("vsn")
meanSdPlot(assay(ntd))
```
```{r}
meanSdPlot(assay(vsd))
```
(12) Heatmap of the count matrix. The heatmap shows the results of 139 donors and 20 genes. If you need to view specific genes, you can change [1:20].
```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("donor_type","donor_sex","donor_age_at_diagnosis")])
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df,show_colnames = FALSE)

```
(13) Heatmap of the sample-to-sample distances
```{r}
sampleDists <- dist(t(assay(vsd)))
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vsd$condition, vsd$type, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```

(14) Principal component plot of the samples (PCA plot)
PCA plot 1 (donor type and donor sex)
```{r}
pcaData <- plotPCA(vsd, intgroup=c("donor_type", "donor_sex"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=donor_sex, shape=donor_type)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
```
PCA plot 2 (donor type and donor age at diagnosis)
```{r}
pcaData <- plotPCA(vsd, intgroup=c("donor_type", "donor_age_at_diagnosis"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=donor_age_at_diagnosis, shape=donor_type)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
```

