---
title: "R Notebook"
output: html_notebook
---

Step 3:
Using data in step 1 to analyse difference between two donor sex (male and female). Draw MA plot, plot counts, variance plot, heatmap, sample-to-sample heatmap, and PCA plot.

(1)load data in R for the DESeq2.
```{r}
rm(list = ls())
counts0<-read.csv("counts.csv",row.names = 1)
donor0<-read.csv("donor.csv",row.names = 1)
```
(2) Donor information is categorized by gender, and Htseq-counts are corresponded to donor information.
```{r}
donor0<- donor0[order(donor0$donor_sex),]
counts1<- t(counts0)
counts1<- cbind(rownames(counts1), data.frame(counts1, row.names=NULL))
counts1<- dplyr::rename(counts1, "icgc_donor_id" = "rownames(counts1)")
```
```{r}
counts2<- dplyr::select(donor0,c(icgc_donor_id))
counts2<- dplyr::inner_join(counts2,counts1,by='icgc_donor_id')
row.names(counts2) <- counts2[, 1]
counts2<- counts2[, -1]
counts2<- t(counts2)
colnames(counts2) == donor0$icgc_donor_id
```
(3)load data in DESeq2 (donor sex).
```{r}
library("DESeq2")
dds<-DESeqDataSetFromMatrix(countData = counts2,
                            colData = donor0,
                            design = ~donor_sex)
```
(4)Pre-filtering: keep the data have more then 10 values in a raw.
```{r}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```
(5) Note on factor levels.
```{r}
dds$donor_type <- factor(dds$donor_type, levels = c("male","female"))
```
(6) Differential expression analysis and save data
```{r}
dds<-DESeq(dds)
res <- results(dds, alpha=0.05)
resOrdered <- res[order(res$pvalue),]
write.csv(as.data.frame(resOrdered), 
          file="donor_sex_male_vs_female_p_order.csv")
```
```{r}
summary(res)
sum(res$padj < 0.05, na.rm=TRUE)
```
(7)Log fold change shrinkage for visualization and ranking by LFC estimates.
```{r}
resultsNames(dds)
resLFC <- lfcShrink(dds, coef="donor_sex_male_vs_female", type="apeglm")
summary(resLFC)
sum(resLFC$padj < 0.05, na.rm=TRUE)
```
(8) draw MS plot with apeglm, normal, and ashr.
```{r}
plotMA(res,  xlim=c(0.01,1e5), ylim=c(-10,15))
resNorm <- lfcShrink(dds, coef=2, type="normal")
resAsh <- lfcShrink(dds, coef=2, type="ashr")
par(mfrow=c(1,3), mar=c(4,4,2,1))
xlim <- c(0.1,1e6); ylim <- c(-1.5,1.5)
plotMA(resLFC, xlim=xlim, ylim=ylim, main="apeglm")
plotMA(resNorm, xlim=xlim, ylim=ylim, main="normal")
plotMA(resAsh, xlim=xlim, ylim=ylim, main="ashr")
```
(9)draw MA-plot with ashr because it retains more functions and is suitable for handling large amounts of data.
```{r}
plotMA(resAsh,xlim=c(0.1,1e6), ylim=c(-6,6))
```
(10) Draw Plot counts with dds. Here I specify the gene which had the smallest p value from the results table created above. You can select the gene to plot by row name or by numeric index.
```{r}
d <- plotCounts(dds, gene=which.min(res$padj), intgroup="donor_sex", 
                returnData=TRUE)
library("ggplot2")
ggplot(d, aes(x=donor_sex, y=count)) + 
  geom_point(position=position_jitter(w=0.1,h=0)) + 
  scale_y_log10(breaks=c(50,500,10000))
```
(11)Extracting transformed values
```{r}
vsd <- vst(dds, blind=FALSE)
ntd <- normTransform(dds)
```
(12) Effects of transformations on the variance
```{r}
library("vsn")
meanSdPlot(assay(ntd))
```
```{r}
meanSdPlot(assay(vsd))
```

(13) Heatmap of the count matrix. The heatmap shows the results of 139 donors and 20 genes. If you need to view specific genes, you can change [626:645].
```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[626:645]
df <- dplyr::select(donor0,c(icgc_donor_id,donor_sex))
rownames(df) = df[,1]
df<- dplyr::select(df,-c(icgc_donor_id))
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=T,
         cluster_cols=FALSE, annotation_col=df, show_colnames = FALSE)

```
(14) Heatmap of the sample-to-sample distances
```{r}
sampleDists <- dist(t(assay(vsd)))
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vsd$condition, vsd$type, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```
(15) Principal component plot of the samples (PCA plot)
```{r}
pcaData <- plotPCA(vsd, intgroup=c("donor_sex"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=donor_sex)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
```





