---
title: "R Notebook"
output: html_notebook
---

Preparation step：
The following packages need to be installed before running the following code. If you have already installed these packages, you can skip this step. 

(1)install packages ggplot2 in R
```{r}
install.packages("ggplot2")
```
(2)install packages DESeq2, org.Hs.eg.db, apeglm, IHW, and vsn in R
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("org.Hs.eg.db")
BiocManager::install("apeglm")
BiocManager::install("IHW")
BiocManager::install("vsn")
```
(3)install packages ashr in R
```{r}
install.packages("devtools")
install_github("stephens999/ashr")
```
(4)install packages pheatmap in R
```{r}
install.packages("pheatmap")
```




Step 1:
Change the downloaded data format to a format suitable for DESesq2 analysis. Change gene_id to symbols.

(1) load Htseq-counts data in R, and keep the required data. Change data format (length data to width data).
```{r}
rm(list = ls())
pasAnno <- read.table(file = 'seq.tsv', sep = '\t', header = TRUE) 
pasAnno1<-subset(pasAnno,select = c(1,8,10))
require(tidyverse)
dat <-  pasAnno1 %>% 
  group_by(icgc_donor_id) %>%
  mutate(index = row_number()) %>%
  pivot_wider(names_from = icgc_donor_id, 
              values_from = raw_read_count) %>%
  dplyr::select(-index)
pasAnno2<-dat
head(pasAnno)
head(pasAnno1)
head(pasAnno2)
```
(2)Delete invalid data (Delete all NULL rows, incomplete data, and all zero rows.)：
```{r}
pasAnno3<- filter(pasAnno2, pasAnno2$DO51966 !='NULL') 
pasAnno3<- dplyr::select(pasAnno3,-c(DO51973,DO6940,DO6635,DO6630,DO6625,DO6620,DO6730,DO6408,DO6406,DO6400,DO6454,DO6450,DO6452,DO6448,DO6444,DO6438,DO6430,DO6426,DO6422,DO6424,DO6498,DO6494,DO6488,DO6484,DO6486,DO6480,DO6482,DO6476,DO6478,DO6472,DO6474,DO6466,DO6464,DO6460,DO6531,DO6534,DO6525,DO6519,DO6516,DO6510,DO6507,DO6504,DO6501,DO6565,DO6561,DO6558,DO6555,DO6552,DO6546,DO6543,DO6585,DO6615,DO6374,DO6360,DO6356,DO6351,DO6396,DO6394,DO6390,DO7108,DO7136,DO7112,DO7048,DO7084,DO6760,DO6418))
pasAnno4<-as.data.frame(pasAnno3)
row.names(pasAnno4)<-pasAnno4[,1]
pasAnno4<-pasAnno4[,-1]
pasAnno4 = pasAnno4[rowSums(pasAnno4) !=0,]
head(pasAnno4)
```
(3)change gene_id to symbol
```{r}
pasAnno5<- cbind(rownames(pasAnno4), data.frame(pasAnno4, row.names=NULL))
colnames(pasAnno5)[1] = 'gene_id'
pasAnno5<-separate(pasAnno5,col=gene_id,into=c("ensembl_id","b"),sep="[.]" )
pasAnno5<- dplyr::select(pasAnno5,-c(b))
library(org.Hs.eg.db)
#ls("package:org.Hs.eg.db")
g2s<-toTable(org.Hs.egSYMBOL)
g2e<-toTable(org.Hs.egENSEMBL)
b<-merge(pasAnno5,g2e,by='ensembl_id',all.x=T)
d<-merge(b,g2s,by='gene_id',all.x=T)
d<-d[,c(142,2:141,1)]
detach("package:org.Hs.eg.db");
pasAnno6<- dplyr::select(d,-c(ensembl_id,gene_id))
require(tidyverse)
pasAnno6<- filter(pasAnno6, pasAnno6$symbol !='na') %>% 
dplyr::distinct(symbol, .keep_all = TRUE)
head(pasAnno6)
```
(4) Store the processed file for subsequent steps.
```{r}
write.csv(pasAnno6, file = "counts.csv", row.names = F, quote = F)
```


